New insights into the Weddell Sea ecosystem applying a network approach

### 2022-12-15 Manuscript/Manuscript.Rmd
Finished Intro

\introduction[Introduction]

The objective of this work was to improve the knowledge on how the Weddell Sea and the species therein may respond to perturbations from ongoing environmental changes. To achieve this we: 1) estimated the strength for each interaction in the Weddell Sea food web, 2) characterised species considering weighted and unweighted properties, and 3) analysed the species’ role in the stability of the food web.

>>>>

# Introduction

<!-- Intro & advantages of studying food webs -->
A network of feeding interactions is an important framework for understanding ecological community structure and a valuable tool for conserving biodiversity and ecosystem functioning through management [@Thompson2012]. Although topological food web analyses considering presence/absence of predator-prey interactions provide important insights into the structure and functioning of ecological communities [e.g. @Kortsch2015; @Marina2018; @Cordone2020; @Rodriguez2022], more information on the nature of an ecological interaction is needed to effectively charaterise ecosystem dynamics and stability. In this regard, quantifying the strength of species interactions is of paramount importance [@Carrara2015; @Allesina2015; @Nilsson2016].

<!-- Interaction strength -->
Several methodologies have been applied to estimate the interaction strength in food webs, where the quantity and quality of data mostly determine which approach is more convenient [@Berlow2004]. Experimental methods combined with dynamic modelling [@Emmerson2004; @Carrara2015], measurements of species abundances trough time [@Fahimipour2014; @Chang2021], and/or estimation of metabolic rates and biomass of all species in the community [@Neutel2014]; all these methods require a large experimental and/or sampling effort that restrict them to simple networks (circa 10 species). Other methods based on predator-prey body mass allometric relationships [@Berlow2009; @Pawar2012] could be applied to larger networks due to less data requirements, although they might have less power with increasingly complex networks unless more traits are included in the estimations [@Jonsson2018]. For food webs composed by nearly 1000 species and about 20000 interactions only those methods with less data requirements are feasible. On of these is the one proposed by @Pawar2012, which allows to obtain interaction strength estimations by combining data on body mass (of consumer -predator- and resource -prey-) and interaction dimensionality (classified as 2 or 3 dimensions).

<!-- The Weddell Sea represents the southerly part of the Atlantic Sector of the Southern Ocean. About one quarter of the Weddell Sea’s entire marine area lies over the continental shelf, stretching along the eastern contour of the Antarctic Peninsula and the Antarctic continent up to 20◦E. The ecosystem of the Weddell Sea is one of the most pristine and relatively unexplored hotspots of marine biodiversity in Antarctica [@Teschke2021]. It contains several areas of ecological significance for birds and marine mammals, such as important breeding and foraging areas [@Hindell2020]. Furthermore, the sea floor is home to large and unique forest-like sponge and filter-feeder assemblages, which are particularly numerous on the continental shelf in the eastern part of the Weddell Sea [@Pineda-Metz2019]. These species assemblages create a structurally complex and diverse habitat for thousands of species [@Barthel1992; @Brey1994; Brandt2007]. -->

<!-- Food web studies about Weddell Sea / Key species or species role -->
The complexity of the network of feeding interactions in the Weddell Sea ecosystem, particularly in the high continental shelf, was introduced by @Jacob2011. It's important to note that this study is the only one addressing the complexity, structure and stability of the Weddell Sea's food web. Few other studies have considered this food web though in a meta-analysis context [@Brose2019; @Perkins2022; @Saravia2022]. The network analysed by @Jacob2011 consisted of 488 species and 16200 predator-prey interactions. Studying the relationships between trophic structure, body size and network structure, the authors suggested that the diversity of predatory species types is essential to better comprehend the Weddell Sea ecosystem. In an attempt to tackle food web stability they performed extinction simulations and showed that the removal of small to medium-sized, and not large, organisms caused a cascade of secondary extinctions. Overall, this suggests the relative importance of predators, rather than prey, in the architecture, functioning and stability of the Weddell Sea food web, which coincides with recent meta-analyses in complex food webs [@Brose2019; @Perkins2022].

<!-- Climate change in Weddell Sea -->
Because of the ice cover and ocean currents, the Weddell Sea is expected to be one of the last regions of the Southern Ocean where the consequences of climate change will manifest [@Teschke2021]. In this regard, the Weddell Sea region might be less affected by sea surface warming compared with other areas of the Southern Ocean. Until then, this region may serve as a retreat for those Antarctic species which either directly depend on sea ice (e.g. krill, emperor penguin, Weddell seal) or have developed such strong adaptations to the polar temperatures during the past millennia that their heat tolerance is low (e.g. most notothenioid fishes). If a climate change scenario is considered, the Weddell Sea continental shelf will be probably a sink area and a refuge for highly cold-adapted species [@Griffiths2017]. While the essential large-scale hydrodynamic relationships are relatively well known for this region [@deSteur2019], information on current distribution, abundance and sensitivity to climate change in the Weddell Sea is only partially available for a few species [@Houstin2022].

In the past decades, network science has become a powerful interdisciplinary approach applied across various spatial scales in polar regions [@Carscallen2012; @Santana2013]. Particularly in the Southern Ocean, important insights have been gained into mechanisms for energy flow, the relative importance and traits of individual species and the influence of environmental variables (e.g., sea-ice) on the structure of local food webs [@Cordone2020; @Rossi2019].

<!-- Objective -->
The objective of this work is to improve the knowledge on the network of feeding interactions and the species' role in the stability of the Weddell Sea ecosystem. To achieve this we: 1) estimated the strength for each interaction in the Weddell Sea food web, 2) characterised species considering weighted and unweighted properties, and 3) analysed the species’ role in the stability of the food web.
